Investigating the Underprovision of Public Goods: Does Segregation Matter?

### 2020-05-05 final_draft_extension_fs.Rmd
cutoff analysis

```{r fp_imputations_original, include = FALSE, eval = FALSE}

@@

fin <- complete(imp_fs) %>%
  filter(!(is.na(dgepercap_cpi)), !(is.na(H_citytract_NHW_i)), !(is.na(diversityinterp)),
         !(is.na(pctblkpopinterp)), !(is.na(pctasianpopinterp)), !(is.na(pctlatinopopinterp)),
         !(is.na(medinc_cpi)), !(is.na(pctlocalgovworker_100)), !(is.na(pctrentersinterp)),
         !(is.na(pctover65)), !(is.na(pctcollegegradinterp)), !(is.na(logpop)))

@@

kable(imp_felm1_sum_table_vars, "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita City Expenditures: Direct General Expenditure per Capita}",
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(general = c("This table shows the results of Trounstine's first model for the effect of segregation on per capita",
  "direct public expenditure using the two-group Theil's H Segregation Index using the new, multiply imputed datasets.",
  "The results present a stark contrast to Trounstine's original results: the main coefficient on segregation index",
  "has decreased to essentially zero and has become statistically insignificant. The very large P-Value may indicate there is",
  "a larger difference between the two models. The coefficient on diversity has increased in magnitude and become",
  "statistically significant."),
           general_title = "Table 4:") %>%
  kable_styling(full_width = TRUE)

@@

```{r results_printing_fs_impute_regressions_3, results = "asis", echo=FALSE, eval = FALSE, invlude = FALSE}

@@

kable(imp_over_baqir_1_sum_table_vars, "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita City Expenditures: Direct General Expenditure per Capita}",
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(general = c("This table shows the results of Trounstine's third model for the effect of segregation on political polarization",
  "(racial vote divide) using the two-group Theil's H segregation Index and including controls for ideology",
  "using the new, multiply imputed datasets. Compared to Trounstine's original results, the main coefficient on",
  "segregation index has decreased in magnitude although, it remains statistically significant. The other",
  "coefficients except for the biracial indicator are insignificant, and the coefficients on percent Asian and",
  "Black have switched signs. Standard errors have also reduced slightly in these specifications compared to",
  "the original analysis in Trounstine (2016)."),
           general_title = "Table 6:") %>%
  kable_styling(full_width = TRUE)

@@

```{r city_pops_histogram, echo = FALSE}

pop_hist_data %>%
  ggplot() +
  geom_histogram(aes(x = log_pop, fill = Analysis)) +
  geom_vline(aes(xintercept = median_log_pop), color = "red") +
  facet_wrap(~Analysis) +
  coord_cartesian(xlim = c(0,15)) +
  ggtitle("Histograms of Log Populaiton of City Observations Included in the Various\nRegression Analyses") +
  labs(caption = "This figure makes clear the differences in city size (in terms of population) among the different analyses. It is\nclear that the listwise deletion of observations with missing values resulted in a sample that almost entirely\nconsisted of larger cities in Trounstine (2016). The distribution of city population in Trounstine (2016) more\nclosely resembles the distribution of the original sample with a greater than 25,000 population cutoff imposed\nthan the original sample. The red lines indicate the median log population in each sample.") +
  ylab("Count") +
  xlab("Log Population") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0))

@@

```{r calc_coeff_sds_original_fs_impute, include = FALSE, eval = FALSE}

imp_felm1_sum[1,2]
imp_under_baqir_1_sum
imp_over_baqir_1_sum
coef(felm1)[1]

summary(felm1)$coeff[2,2]

model_coeff <- tibble(
  variable = c("H_citytract_NHW_i", "diversityinterp", 
               "H_citytract_NHW_i", "diversityinterp", 
               "H_citytract_NHW_i", "diversityinterp",
               "H_citytract_NHW_i", "diversityinterp"),
  estimate = c(imp_felm1_sum[1,2], imp_felm1_sum[2,2],
               imp_under_baqir_1_sum[1,2], imp_under_baqir_1_sum[2,2],
               imp_over_baqir_1_sum[1,2], imp_over_baqir_1_sum[2,2],
               coef(felm1)[1], coef(felm1)[2]),
  std_error = c(imp_under_baqir_1_sum[1,3], imp_under_baqir_1_sum[2,3],
                imp_felm2_sum[1,3], imp_felm2_sum[2,3],
                imp_over_baqir_1_sum[1,3], imp_over_baqir_1_sum[2,3],
                summary(felm1)$coeff[1,2], summary(felm1)$coeff[2,2]),
  model = c(1, 1,
            2, 2,
            3, 3,
            4, 4)
)

@@

model_coeff %>%
  ggplot(aes(x = model, y = estimate)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - std_error, ymax = estimate + std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(~variable)

>>>>

```{r fp_imputations_original, include = FALSE}

@@

@@

kable(imp_felm1_sum_table_vars, "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for all Cities}",
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(general = c("This table shows the results of Trounstine's first model for the effect of segregation on per capita direct",
  "public expenditure using the two-group Theil's H Segregation Index using the new, multiply imputed datasets.",
  "The results present a stark contrast to Trounstine's original results: the main coefficient on segregation index",
  "has decreased to essentially zero and has become statistically insignificant. The very large P-Value may indicate",
  "there is a larger difference between the two models. The coefficient on diversity has increased in magnitude and",
  "become statistically significant."),
           general_title = "Table 4:") %>%
  kable_styling(full_width = TRUE)

@@

```{r results_printing_fs_impute_regressions_3, results = "asis", echo=FALSE, eval = FALSE, include = FALSE}

@@

kable(imp_over_baqir_1_sum_table_vars, "latex",
      caption = "\\textbf{Effect of Segregation on Overall per Capita Direct General City Expenditures: Imputed Data for Cities with Populations of 25,000+}",
      booktabs = T) %>%
    row_spec(0, bold = TRUE) %>%
  footnote(general = c("This table shows the results of Trounstine's third model for the effect of segregation on political polarization",
  "(racial vote divide) using the two-group Theil's H segregation Index and including controls for ideology",
  "using the new, multiply imputed datasets. Compared to Trounstine's original results, the main coefficient on",
  "segregation index has decreased in magnitude although, it remains statistically significant. The other",
  "coefficients except for the biracial indicator are insignificant, and the coefficients on percent Asian and",
  "Black have switched signs. Standard errors have also reduced slightly in these specifications compared to",
  "the original analysis in Trounstine (2016)."),
           general_title = "Table 6:") %>%
  kable_styling(full_width = TRUE)

@@

```{r city_pops_histogram, echo = FALSE, warning = FALSE}

pop_hist_data %>%
  ggplot() +
  geom_histogram(aes(x = log_pop, fill = Analysis)) +
  geom_vline(aes(xintercept = median_log_pop), color = "red") +
  facet_wrap(~Analysis) +
  coord_cartesian(xlim = c(0,15)) +
  ggtitle("Histograms of Log Populaiton of City Observations Included in the\nVarious Regression Analyses") +
  labs(caption = "This figure makes clear the differences in city size (in terms of population) among the different analyses.\n It is clear that the listwise deletion of observations with missing values resulted in a sample that almost entirely\nconsisted of larger cities in Trounstine (2016). The distribution of city population in Trounstine (2016)\nmore closely resembles the distribution of the original sample with a greater than 25,000 population\ncutoff imposed than the original sample. The red lines indicate the median log population in each sample.") +
  ylab("Count") +
  xlab("Log Population") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.text.x = element_text(size = 8))

@@

```{r top_1_%_pop}

fin_imp_top_1 <- fin_imp %>%
  filter(logpop > quantile(fin_seg$logpop, c(.95), na.rm = T))

nrow(fin_imp_top_1)

ini <- mice(fin_imp_top_1, maxit = 0, print = FALSE)

imp_fs_pop_top_1 <- mice(fin_imp_top_1, method = meth, maxit = 5)

fit_fs_pop_top_1 <- with(imp_fs_pop_top_1, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

pool_fs_pop_top_1 <- pool(fit_fs_pop_top_1)
imp_fs_pop_top_1_sum <- summary(pool_fs_pop_top_1)

```{r}
fin_seg %>%
  ggplot(aes(x = logpop)) +
  geom_histogram() +
  geom_vline(aes(xintercept = log(25000)))

log(25000)
```

```{r bottom_60_%_pop}

quantile(fin_seg$logpop, c(.60), na.rm = T)

fin_imp_bottom_60 <- fin_imp %>%
  filter(logpop < quantile(fin_seg$logpop, c(.60), na.rm = T))

nrow(fin_imp_bottom_60)

ini <- mice(fin_imp_bottom_60, maxit = 0, print = FALSE)

imp_bottom_60 <- mice(fin_imp_bottom_60, method = meth, maxit = 5)

fit_imp_bottom_60 <- with(imp_bottom_60, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

pool_fs_bottom_60 <- pool(fit_imp_bottom_60)
imp_fs_pop_bottom_60<- summary(pool_fs_bottom_60)

```{r 50k}

fin_imp_50k <- fin_imp %>%
  filter(logpop < log(50000))

nrow(fin_imp_50k)

ini <- mice(fin_imp_50k, maxit = 0, print = FALSE)

imp_50k <- mice(fin_imp_50k, method = meth, maxit = 1)

fit_imp_50k <- with(imp_50k, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

# pool the analyses: insignificant for both

pool_fs_50k<- pool(fit_imp_50k)
imp_fs_pop_50k<- summary(pool_fs_50k)

```{r 250k}

fin_imp_250k <- fin_imp %>%
  filter(logpop < log(250000))

nrow(fin_imp_250k)

ini <- mice(fin_imp_250k, maxit = 0, print = FALSE)

imp_250k <- mice(fin_imp_250k, method = meth, maxit = 1)

fit_imp_250k <- with(imp_250k, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

pool_fs_250k<- pool(fit_imp_250k)
imp_fs_pop_250k<- summary(pool_fs_250k)

```{r 500k}

fin_imp_250_500k <- fin_imp %>%
  filter(logpop > log(250000) & logpop < log(500000))

nrow(fin_imp_250k)

ini <- mice(fin_imp_250k, maxit = 0, print = FALSE)

imp_250k <- mice(fin_imp_250k, method = meth, maxit = 1)

fit_imp_250k <- with(imp_250k, felm(dgepercap_cpi ~ H_citytract_NHW_i + diversityinterp + pctblkpopinterp + pctasianpopinterp + pctlatinopopinterp + medinc_cpi + pctlocalgovworker_100 + pctrentersinterp + pctover65 + pctcollegegradinterp + logpop | factor(geo_id2) |0| geo_id2))

pool_fs_250k<- pool(fit_imp_250k)
imp_fs_pop_250k<- summary(pool_fs_250k)

```{r calc_coeff_sds_original_fs_impute, include = FALSE}

imp_felm1_sum[1,2]
imp_under_baqir_1_sum
imp_over_baqir_1_sum
#coef(felm1)[1]
imp_fs_pop_top_1_sum

#summary(felm1)$coeff[2,2]

# lets put the models in order of increasing median population size in the sample.

pop_hist_data %>%
  select(Analysis, median_log_pop) %>%
  arrange(median_log_pop) %>%
  unique()

model_coeff <- tibble(
  variable = c("White/Non-White Seg. Index", "Diversity", 
               "White/Non-White Seg. Index", "Diversity", 
               #"White/Non-White Seg. Index", "Diversity",
               "White/Non-White Seg. Index", "Diversity",
               "White/Non-White Seg. Index", "Diversity"),
  estimate = c(imp_under_baqir_1_sum[1,2], imp_under_baqir_1_sum[2,2],
               imp_felm1_sum[1,2], imp_felm1_sum[2,2],
               #coef(felm1)[1], coef(felm1)[2],
               imp_over_baqir_1_sum[1,2], imp_over_baqir_1_sum[2,2],
               imp_fs_pop_top_1_sum[1,2], imp_fs_pop_top_1_sum[2,2]),
  std_error = c(imp_under_baqir_1_sum[1,3], imp_under_baqir_1_sum[2,3],
                imp_felm1_sum[1,3], imp_felm1_sum[2,3],
                #summary(felm1)$coeff[1,2], summary(felm1)$coeff[2,2],
                imp_over_baqir_1_sum[1,3], imp_over_baqir_1_sum[2,3],
                imp_fs_pop_top_1_sum[1,3], imp_fs_pop_top_1_sum[2,3]),
  model = c(1, 1, 
            2, 2,
            #3, 3,
            3, 3,
            5, 5)
)

@@

```{r coef_plot, echo = FALSE}

coef.labels <- c("Diversity", "White/Non-White Seg. Index")
names(coef.labels) <- c("Diversity", "White/Non-White Seg. Index")

model_coeff %>%
  ggplot(aes(x = model, y = estimate)) +
  geom_point()+
  geom_errorbar(aes(ymin = estimate - 2 * std_error, ymax = estimate + 2 * std_error, width = .2)) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(~variable, labeller = labeller(variable = coef.labels)) +
  ggtitle("Coefficient Plots for Diversity and White/Non-White Segregation Indices Across\nthe Models")+
  xlab("Models Ordered by Increasing Median Population Size of the Sample") +
  ylab("Estimate and 95% Confidence Interval") +
  labs(caption = "This plot demonstrates how the significance of diversity and segregation in predicting public\ngoods spending varies with city size. As city populaiton size increases, diversity tends to have\nless explanatory power for public goods spending while segregation has more. The opposite\nis true when considering city population size decreasing: segregation becomes less important\nas diversity becomes increasingly important.") +
  theme_bw()+
  theme(plot.caption = element_text(hjust = 0))
